MSDS6372Project1

### 2021-02-02 MSDS6372Project1.Rmd
upload tree regression and update test RMSE

# LASSO and STEPWISE came to slight different Model in terms of predictor and r-squared value.
# LASSO r-squared 0.8426, adjusted r-squared 0.8419
# Stepwise r-squared 0.843, adjusted r-squared 0.8424
#compare test ASE results between stepwise model and LASSO
test = test[,-(1:2)]
testMSRP = test$MSRP # GET real result 
test = test[,-13] # Delete MSRP from test 
# Calculate stepwise test ASE first
pred.test.step = predict(log.model8.step,test)
# exp values to return to original value
pred.test.step = exp(pred.test.step)
test_ASE_STEP = mean((testMSRP-pred.test.step)^2)
test_ASE_STEP #571362858

#Calculate LASSO test ASE
pred.test.lasso = predict(log.model7,test)
pred.test.lasso = exp(pred.test.lasso)
test_ASE_LASSO = mean((testMSRP-pred.test.lasso)^2)
test_ASE_LASSO #571362858

@@

# LogMSRP with Second Degree EngineHP picked by LASSO test result
lasso.log.pred1 = predict(lasso.lm.log,test)
# Recover the value by exp
lasso.log.pred1 = exp(lasso.log.pred1)
testASE_LASSO_LOG = mean((test.result-lasso.log.pred1)^2)
testASE_LASSO_LOG #24577120

@@

>>>>

# LASSO and STEPWISE came to slight different Model in terms of predictor and r-squared value.
# LASSO r-squared 0.8426, adjusted r-squared 0.8419
# Stepwise r-squared 0.843, adjusted r-squared 0.8424
#compare test ASE results between stepwise model and LASSO
test = test[,-(1:2)]
testMSRP = test$MSRP # GET real result 
test = test[,-13] # Delete MSRP from test 
# Calculate stepwise test ASE first
pred.test.step = predict(log.model8.step,test)
# exp values to return to original value
pred.test.step = exp(pred.test.step)
test_ASE_STEP = mean((testMSRP-pred.test.step)^2)
sqrt(test_ASE_STEP) #test RMSE = 23903.2

#Calculate LASSO test ASE
pred.test.lasso = predict(log.model7,test)
pred.test.lasso = exp(pred.test.lasso)
test_ASE_LASSO = mean((testMSRP-pred.test.lasso)^2)
sqrt(test_ASE_LASSO) ##test RMSE = 23903.2

@@

# LogMSRP with Second Degree EngineHP picked by LASSO test result
lasso.log.pred1 = predict(lasso.lm.log,test)
# Recover the value by exp
lasso.log.pred1 = exp(lasso.log.pred1)
testASE_LASSO_LOG = mean((test.result-lasso.log.pred1)^2)
sqrt(testASE_LASSO_LOG) #test RMSE = 4957.532

@@

# Note: Rerun the code chunk for clean train/test/validate datasets
#### Objective 2 KNN Regression
KNN Regression is a non-parametric machine learning approach, it first identifies the K training observations that are closest to X0, it then estimates the response using the average of all the training response identified. The optimal value for K will depend on the bias-variance trade off. A small K provides the most flexible fit, which will have low bias but high variance. This variance is due to the fact that the prediction in a given region is entirely dependent on just one observation. In contrast, larger values of K provide a smoother and less variable fit; the prediction in a region is an average of several points, and so changing one observation has smaller effect. However, the smoothing may cause bias by masking some of the structure in the response. -- (ISLR Seventh Edition) Note: KNN regression can only handles numerical variables.

```{r}
library(caret)
############################################################
#testing KNN regression - only handles numeric variables####
############################################################

#getting training response
trainY = train$MSRP
#getting training predictors
trainX = train[,-c(1,2,4,7,8,10,11,15,16,17,18,19,20,21,22,23,24,25,26)]
# Doing the same for test
testY = test$MSRP
testX = test[,-c(1,2,4,7,8,10,11,15,16,17,18,19,20,21,22,23,24,25,26)]

fit = knnreg(trainX,trainY,k=3)
plot(testY,predict(fit,testX))
#Try testing testASE
knn.k3.pred = predict(fit,testX)
testASE_KNN_k3 = mean((testY-knn.k3.pred)^2)
sqrt(testASE_KNN_k3)

############# explore optimal k using test ASE ###############
iterations = 100
testASE_List = c()
for(j in 1:iterations){
  fit = knnreg(trainX,trainY,k=j)
  knn.pred = predict(fit,testX)
  testASE_loop = mean((testY-knn.pred)^2)
  testASE_List[j] = sqrt(testASE_loop)
}
par(mfrow=c(1,1))
plot(1:iterations,testASE_List,type="l",xlab="# of k",ylab="test Root Mean Square Error")
index<-which(testASE_List==min(testASE_List))
index

# seems k=2 gives the best result according to our test ASE
fit = knnreg(trainX,trainY,k=2)
plot(testY,predict(fit,testX))
#Try testing testASE
knn.k2.pred = predict(fit,testX)
testASE_KNN_k2 = mean((testY-knn.k2.pred)^2)
sqrt(testASE_KNN_k2)
